Extracting The Power of Dependent Types

### 2021-07-15 paper.tex
Shrink to fit 12 pages :)
Use proper color for macro calls
Add section references to contributions

\usepackage{agda}
\usepackage{newunicodechar}
\usepackage[final]{listings}

@@

\newunicodechar{←}{\ensuremath{\leftarrow}}
\newunicodechar{×}{\ensuremath{\times}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{Π}{\ensuremath{\Pi}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≅}{\ensuremath{\cong}}
\newunicodechar{≐}{\ensuremath{\doteq}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∧}{\ensuremath{\land}}
\newunicodechar{∨}{\ensuremath{\lor}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{⊔}{\ensuremath{\sqcup}}
\newunicodechar{∷}{\ensuremath{{::}}}
\newunicodechar{ℓ}{\ensuremath{\ell}}
\newunicodechar{₀}{\ensuremath{{_0}}}
\newunicodechar{₁}{\ensuremath{{_1}}}
\newunicodechar{₂}{\ensuremath{{_2}}}
\newunicodechar{₃}{\ensuremath{{_3}}}
\newunicodechar{₄}{\ensuremath{{_4}}}
\newunicodechar{₅}{\ensuremath{{_5}}}
\newunicodechar{₆}{\ensuremath{{_6}}}
\newunicodechar{₇}{\ensuremath{{_7}}}
\newunicodechar{₈}{\ensuremath{{_8}}}
\newunicodechar{₉}{\ensuremath{{_9}}}
\newunicodechar{⟨}{\ensuremath{{\langle}}}
\newunicodechar{⟩}{\ensuremath{{\rangle}}}
\newunicodechar{̧}{\c}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{∸}{̣\ensuremath{-}}
\newunicodechar{ᵇ}{\ensuremath{{}^{b}}}

% Some shortcut commands for agda symbols
\newcommand{\AD}[1]{\AgdaDatatype{#1}}
\newcommand{\AC}[1]{\AgdaInductiveConstructor{#1}}
\newcommand{\AF}[1]{\AgdaFunction{#1}}
\newcommand{\AB}[1]{\AgdaBound{#1}}
\newcommand{\AK}[1]{\AgdaKeyword{#1}}
\newcommand{\AR}[1]{\AgdaField{#1}}
\newcommand{\AM}[1]{\AgdaModule{#1}}
\newcommand{\AN}[1]{\AgdaNumber{#1}}
\newcommand{\AS}[1]{\AgdaString{#1}}

@@

@@

\item We demonstrate this approach by a concrete example: an embedding
  of (a subset of the) PostScript language into Agda.

\item We give several examples of PostScript programs written in our
  embedding, and demonstrate how to statically enforce properties such
  as the absence of stack underflows, termination, and functional
  correctness.

\item We show how to use reflection in Agda to implement an extractor
  that translates code written in our embedding to plain PostScript.

\item We show how to our approach provides easy support for partial
  evaluation by making use of normalisation in the host language. This
  allows us to use functions of the host language as macros, as well
  as automatically apply optimizations through the use of rewrite
  rules.
\end{itemize}

The rest of this paper is structured as follows. \secref{background}
provides background information on Agda and its reflection
capabilities. \secref{embedding} describes our embedding of PostScript
in Agda, and gives several examples of programs written in
it. \secref{extraction} shows how to implement an extractor for this
embedding using reflection. Next, \secref{partial-evaluation}
demonstrates how to apply partial evaluation to programs prior to
extraction by using Agda's normalisation procedure. Finally,
\secref{related} discusses related work and \secref{conclusion}
concludes.

>>>>

\usepackage{agda}
\usepackage{newunicodechar}
\usepackage[final]{listings}
\lstset{
    basicstyle=\small,
}

@@

\newunicodechar{←}{\ensuremath{\leftarrow}}
\newunicodechar{×}{\ensuremath{\times}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{Π}{\ensuremath{\Pi}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≅}{\ensuremath{\cong}}
\newunicodechar{≐}{\ensuremath{\doteq}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∧}{\ensuremath{\land}}
\newunicodechar{∨}{\ensuremath{\lor}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{⊔}{\ensuremath{\sqcup}}
\newunicodechar{∷}{\ensuremath{{::}}}
\newunicodechar{ℓ}{\ensuremath{\ell}}
\newunicodechar{₀}{\ensuremath{{_0}}}
\newunicodechar{₁}{\ensuremath{{_1}}}
\newunicodechar{₂}{\ensuremath{{_2}}}
\newunicodechar{₃}{\ensuremath{{_3}}}
\newunicodechar{₄}{\ensuremath{{_4}}}
\newunicodechar{₅}{\ensuremath{{_5}}}
\newunicodechar{₆}{\ensuremath{{_6}}}
\newunicodechar{₇}{\ensuremath{{_7}}}
\newunicodechar{₈}{\ensuremath{{_8}}}
\newunicodechar{₉}{\ensuremath{{_9}}}
\newunicodechar{⟨}{\ensuremath{{\langle}}}
\newunicodechar{⟩}{\ensuremath{{\rangle}}}
\newunicodechar{̧}{\c}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{∸}{̣\ensuremath{-}}
\newunicodechar{ᵇ}{\ensuremath{{}^{b}}}
\newunicodechar{✔}{\checkmark}

% Some shortcut commands for agda symbols
\newcommand{\AD}[1]{\AgdaDatatype{#1}}
\newcommand{\AC}[1]{\AgdaInductiveConstructor{#1}}
\newcommand{\AF}[1]{\AgdaFunction{#1}}
\newcommand{\AB}[1]{\AgdaBound{#1}}
\newcommand{\AK}[1]{\AgdaKeyword{#1}}
\newcommand{\AR}[1]{\AgdaField{#1}}
\newcommand{\AM}[1]{\AgdaModule{#1}}
\newcommand{\AN}[1]{\AgdaNumber{#1}}
\newcommand{\AS}[1]{\AgdaString{#1}}
\newcommand{\AMA}[1]{\AgdaMacro{#1}}

@@

\renewcommand{\AgdaCommentFontStyle}[1]{\textsf{#1}}

@@

\item We demonstrate this approach by a concrete example: an embedding
  of (a subset of the) PostScript language into Agda
  (\secref{embedding}). We give several examples of PostScript
  programs written in our embedding, and demonstrate how to statically
  enforce properties such as the absence of stack underflows,
  termination, and functional correctness.

\item We show how to use reflection in Agda to implement an extractor
  that translates code written in our embedding to plain PostScript
  (\secref{extraction}).

\item We show how to our approach provides easy support for partial
  evaluation by making use of normalisation in the host language. This
  allows us to use functions of the host language as macros, as well
  as automatically apply optimizations through the use of rewrite
  rules (\secref{partial-evaluation}).
\end{itemize}

\begin{comment}
The rest of this paper is structured as follows. \secref{background}
provides background information on Agda and its reflection
capabilities. \secref{embedding} describes our embedding of PostScript
in Agda, and gives several examples of programs written in
it. \secref{extraction} shows how to implement an extractor for this
embedding using reflection. Next, \secref{partial-evaluation}
demonstrates how to apply partial evaluation to programs prior to
extraction by using Agda's normalisation procedure. Finally,
\secref{related} discusses related work and \secref{conclusion}
concludes.
\end{comment}
